---
title: "tidyverse Overview: Russian Trolls"
author: "Ryan Wesslen"
date: "July 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## Intro

### install and call packages

First, we need to get `tidyverse`. Recall, you only need to install once (but you need to do it for each computer).

After installing, you then need to call the library to "activate" it.

```{r}
# install packages
# install.packages("tidyverse")

# call (i.e., activate) tidyverse
library(tidyverse)
```

## load csv file

Next, let's load the csv file from the url.

```{r}
url <- "http://nodeassets.nbcnews.com/russian-twitter-trolls/tweets.csv"
tweets <- read_csv(url)
```

## Count Function

With the dataframe, let's use the `count()` function to count the number of tweets per user (user_key).

```{r}
# count function
count(tweets, user_key)

# assignment operator
c <- count(tweets, user_key)

# piping
c2 <- tweets %>%
  count(user_key)

# group_by and summarise
c3 <- tweets %>%
  group_by(user_key) %>%
  summarise(Count=n())
```

**Question**: What is the difference between these three dataframes?

## What were the top tweets by retweet and favorite (likes) counts?

Let's use other `dplyr` functions to examine the top retweets...

```{r}
tweets %>%
  arrange(desc(retweet_count)) %>%
  select(user_key, created_str, text, retweet_count)
```

... and most favorited (liked) tweets.

```{r}
tweets %>%
  arrange(desc(favorite_count)) %>%
  select(user_key, created_str, text, favorite_count)
```

We can then use ggplot to count the number of tweets per day.

```{r}
# ggplot
sum <- tweets %>%
  group_by(Date = as.Date(created_str)) %>% # convert string to Date format
  summarise(Count = n())

# create ggplot
ggplot(data = sum, aes(x = Date, y = Count)) +
  geom_line()

tweets %>%
  group_by(Date = as.Date(created_str)) %>%
  summarise(Count = n())

ggplot(data = sum, aes(x = Date, y = Count)) +
  geom_line() +
  geom_point() +
  labs(title = "Daily Tweet Counts by Russian Trolls",
       x = "Time (Day)",
       y = "Number of Tweets",
       caption = "Source: NBC News")
```

We can also save the plot to an object (m).

```{r}
# saving as an object
m <- ggplot(data = sum, aes(x = Date, y = Count)) +
  geom_line() +
  geom_point()

# you can manually call the object
m
```

With that object, you can add new layers to the plot.

```{r}
# you can add layers
m + labs(title = "Daily Tweet Counts by Russian Trolls",
         x = "Time (Day)",
         y = "Number of Tweets",
         caption = "Source: NBC News")
```

## Stringr to filter text

Let's use the `stringr` package to filter tweets that mention either #keithlamontscott or Charlotte (| is the "OR" operator).

```{r}
# use stringr for text filter -- part of the tidyverse 
library(stringr)

term <- "#keithlamontscott|charlotte"

q <- tweets %>%
  filter(str_detect(tolower(tweets$text), tolower(term)))
```

### Add in full pipeline

Let's now link the data

```{r}
# with full pipeline
term <- "obama"

q <- tweets %>%
  filter(str_detect(tolower(tweets$text), term)) %>%
  group_by(Date = as.Date(created_str)) %>%
  summarise(Count=n()) %>%
  ggplot(aes(x = Date, y = Count)) +
  geom_line() +
  labs(title = paste0("Tweets filtered by `",term,"`"),
  x = "Day",
  y = "Number of Tweets")

q
```

## first html widget: plotly

Last, let's use the `plotly` package to make this ggplot interactive.

```{r}
# install.packages("plotly")
library(plotly)

ggplotly(q)
```
