---
title: "Day 1: tidyverse"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr); library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


## ggplot intro

### Call tidyverse 

First call tidyverse to activate it (it's already installed).

```{r library, exercise=TRUE}

```

### Dataset

Next, type in `mpg` to view the mpg dataset. (It's attached to ggplot2)

What do you notice about the data? Note the different formats (e.g., character, numeric, etc.)

```{r mpg, exercise=TRUE, exercise.lines = 3}

```

### ggplot2 simple

Each row is a car by a specific year.

First, use ggplot to display a scatterplot putting `displ` on the x-axis and `hwy` (miles per gallon highway) on the y-axis.

```{r ggplot1, exercise=TRUE, exercise.lines = 3}

```

```{r ggplot1-hint, eval=F}
ggplot(data = <dataframe>) +
  geom_point(mapping = aes(x = <column 1>, y = <column 2>))
```

### Quick exercise

Try these three exercises:

1.  Run ggplot(data = mpg). What do you see?

2.  Make a scatterplot of hwy vs cyl.

3.  What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r exercise1, exercise=TRUE, exercise.lines = 3}

```

## Aesthetics

Let's now consider adding in aesthetic mappings (color, shape, size, etc.)

### Color

To change the color, change the colors of the points by the `class` aesethetic.

```{r aes1, exercise=TRUE, exercise.lines = 3}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

### Size

Next, remove class from color and change it to modify the size.

```{r aes2, exercise=TRUE, exercise.lines = 3}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

### Alpha and Shape

Try to change `size` to `alpha` and then `shape`. What happens?

```{r aes3, exercise=TRUE, exercise.lines = 3}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

### Mappings 

It's important to note the significance of putting aesthetics within the `aes` function.

If we wanted to change the color of all points (as opposed to being set by a variable), we can remove the color parameter to outside of the `aes` function.

Try to run this plot. What happens if you change the color parameter to "red"? Try another color.

```{r aes4, exercise=TRUE, exercise.lines = 3}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

### Exercise 2

What is wrong with this plot? Why are the points not blue?

```{r aes5, exercise=TRUE, exercise.lines = 3}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

## Faceting

Let's consider using the facet functions to add in multiple plots by variable(s).

```{r facet1, exercise=TRUE, exercise.lines=4}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(year ~ class)
```

### Facet Wrap

Next, rerun but use the `facet_wrap()` layer instead of the `facet_grid()`.

```{r facet2, exercise=TRUE, exercise.lines=4}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ manufacturer, nrow = 3)
```

What happens if you change the `nrow` parameter.

## Geom's

### One dimensional 

For a histogram, you can use the function `geom_histogram()`. This geom is only one-dimnesional, so it only requires an x value, not a y value.

Add to this `geom_histogram(aes(x = hwy))`

```{r geom1, exercise=TRUE, exercise.lines=4}
ggplot(mpg)
```

You can now add additional parameters to the plot. For example, add `bins = 50`.

```{r geom2, exercise=TRUE, exercise.lines=4}
ggplot(mpg) + 
  geom_histogram(aes(x = hwy))
```

There are many other one dimensional geom's.

For example, this time add in `geom_density(aes(x = hwy))`.

```{r geom3, exercise=TRUE, exercise.lines=4}
ggplot(mpg)
```

For this plot, we can split it by class by adding the `color = class` aesthetic.

```{r geom4, exercise=TRUE, exercise.lines=4}
ggplot(mpg) +
  geom_density(aes(x = hwy))
```

```{r geom4-hint, eval=F}
Make sure that the parameter is within the aes() function!
```

In this geom, color only means the outline. Instead, use `fill` instead of `color` to fill in the color of the density plots with the hwy field.

```{r geom5, exercise=TRUE, exercise.lines=4}
ggplot(mpg) +
  geom_density(aes(x = hwy))
```

But as you can see, this is too dark. Now add in `alpha = 0.5` to change the transparency of the fill to 0.5 (out of 1).

```{r geom6, exercise=TRUE, exercise.lines=4}
ggplot(mpg) +
  geom_density(aes(x = hwy, fill = class))
```

### Two dimensional

geom_boxplot()

## dplyr and piping

### filter()

Sometimes you will need to do data transformation (i.e., munging) before analyzing data. That's where the `dplyr` package in `tidyverse` will come in handy.

Let's consider some dplyr functions.

First, let's use the `filter()` function to keep only records in 1999.

```{r dplyr, exercise=TRUE}
filter(mpg, year == 1999)
```

Recall, there's different ways to write this function. We can use piping `%>%`.

```{r dplyr2, exercise=TRUE}
mpg %>%
  filter(year == 1999)
```

Notice we get the same result.

Let's now combine this result with piping to ggplot.

```{r dplyr3, exercise=TRUE}
mpg %>%
  filter(year == 1999) %>%
  ggplot() +
  geom_point(aes(x = displ, y = hwy))
```

Notice that ggplot uses a `+` pipe instead of `%>%`. 

### mutate()

First, recall this example but with the modification that year is added for the color.

Why does the color automatically take on a range of blue values?

```{r dplyr4, exercise=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = year)) + 
  facet_wrap(~ class, nrow = 3)
```

What we need to do is reformat year before passing it to ggplot.

To do this, we can use the `mutate()` function in ggplot.

```{r dplyr5, exercise=TRUE}
mpg %>%
  mutate(year = as.factor(year)) %>%
  ggplot() + 
  geom_point(mapping = aes(x = displ, y = hwy, color = year)) + 
  facet_wrap(~ class, nrow = 3)
```

